---
title: "Build MTX predictor"
author: "Tjardo M"
date: "31-01-2023"
output: html_document
---

Initialize working environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/exports/reum/tdmaarseveen/RA_Clustering")

# Set library folder
.libPaths("/exports/reum/tdmaarseveen/Rlibs")
remove.packages("rlang", lib.loc = .libPaths()[x])

setwd("/exports/reum/tdmaarseveen/RA_Clustering")
getwd()
```

## #1: Plot missingness [Unused]

## #1.1 Import lab data

```{r}
library("mice")
library("tidyverse")
library(readr)
library(VIM)

df_lab <- read_delim("new_data/3_wide/DF_Lab_wide_Final_wManRX.csv", "|", escape_double = FALSE, trim_ws = TRUE)

colnames(df_lab) <- c("pseudoId", "MCV", "Ht","Leuko", "MCHC", "MCH",  "BSE", "Hb", "Trom", "aCCP", "RF", 'Age', 'Sex', "Mannequin") # , "Man" , "Rx"
#head(df_lab)

plot_missing(df_lab, "LAB")
```

## #2: Survival analysis

### #2.1: Import data

Import survival data that we generated in notebook 7 (7_survival_analysis.ipynb) at step #7.2.3

```{r}
library("readr")
library(reticulate)

# Initalize conda environment w/ required python packages
#use_condaenv("convae_architecture/envs", required=TRUE)
# offshoots/IMPROVED/df_survivaldata_improved_Remission.csv
# 5_clustering/df_survivaldata_mmae.csv
# Import survival data (generate in #7.2.3 7_survival_analysis.ipynb)
df_survival <- as.data.frame(read_csv("new_data/offshoots/IMPROVED/df_survivaldata_improved_Remission.csv", col_types = cols(...1 = col_skip())))
head(df_survival)
```



### #2.2 Render Kaplan-Meier curve for 1 year
```{r}
library(survival)
library(survminer)
library("ggpubr") #for ggarrange

df_survival$aCCP <- as.integer(df_survival$aCCP)
df_survival$Cluster <- as.factor(df_survival$Cluster)
df_survival$time <- as.integer(df_survival$time) 
df_survival$event_1y_Rem <- as.integer(df_survival$event_1y_Rem) 
#df_survival$event_1y <- as.integer(df_survival$event_1y) 
df_survival$MannTillTreat <- as.integer(df_survival$MannTillTreat) 
```



# Show added effect of clustering on top of ACPA-MTX interaction with an adjusted COX-regression for other covariates 
```{r}
library(survival)

fit1 <- coxph(Surv(time, event_1y) ~ aCCP + RF + Age + Sex , data=df_survival)
fit2 <- coxph(Surv(time, event_1y)~ Cluster + aCCP + RF + Age + Sex , data=df_survival) 

anova(fit1,fit2,test="Chisq")
```
```{r}
fit2 <- coxph(Surv(time, event_1y)~ Cluster + aCCP + RF + Age + Sex , data=df_survival) 
fit2
```

# Visualize Kaplan Meier curves

```{r}


# Cluster+
fitLC  <- survfit(Surv(time, event_1y) ~ Cluster, data = df_survival)

p1 <- ggsurvplot(fitLC
             , data = df_survival
             , risk.table = TRUE
             , conf.int=TRUE
             , pval = TRUE
             , break.x.by = 61
             , break.time.by= 61
             , xlab ="Time in Months"
             # , xlim = c(0, 60)        ## commented out !
             , ylab="% of patients still on initial MTX"
             , ggtheme = theme_minimal()
             , risk.table.y.text.col = T
             , risk.table.y.text = FALSE
             , risk.table.fontsize = 2
              , cumevents = TRUE
              , cumevents.fontsize =2   
             
             )

### save parts of the original graph    
surv_plot <- p1$plot
surv_table <- p1$table# +

### zoom in on the surv_plot    
surv_plot2 <- surv_plot + coord_cartesian(xlim = c(0,366), ylim = c(0.5,1.0))
surv_table <- surv_table + coord_cartesian(xlim = c(0,366))

### put it back together
ggarrange(surv_plot2, surv_table, ncol = 1, heights = c(3, 2))

# Save figure
ggsave("figures/5_treatment/survival_1y.png")
ggsave("figures/5_treatment/survival_1y.pdf")
#surv_plot2
#+ coord_cartesian(xlim=c(0, 365))#  xlim=c(0, 365),

```

### Check for aCCP stratification

```{r}


# Cluster+
fitLC  <- survfit(Surv(time, event_1y) ~ Cluster+aCCP, data = df_survival)

p1 <- ggsurvplot(fitLC
             , data = df_survival
             , risk.table = TRUE
             , conf.int=TRUE
             , pval = TRUE
             , break.x.by = 61
             , break.time.by= 61
             , xlab ="Time in Months"
             # , xlim = c(0, 60)        ## commented out !
             , ylab="% of patients still on initial MTX"
             , ggtheme = theme_minimal()
             , risk.table.y.text.col = T
             , risk.table.y.text = FALSE
             , risk.table.fontsize = 2
              , cumevents = TRUE
              , cumevents.fontsize =2   
             
             )

### save parts of the original graph    
surv_plot <- p1$plot
surv_table <- p1$table# +

### zoom in on the surv_plot    
surv_plot2 <- surv_plot + coord_cartesian(xlim = c(0,366), ylim = c(0.5,1.0))
surv_table <- surv_table + coord_cartesian(xlim = c(0,366))

### put it back together
ggarrange(surv_plot2, surv_table, ncol = 1, heights = c(3, 1))

# Save figure
ggsave("figures/5_treatment/survival_1y_aCCPstrat.png")
ggsave("figures/5_treatment/survival_1y_aCCPstrat.pdf")
#surv_plot2
#+ coord_cartesian(xlim=c(0, 365))#  xlim=c(0, 365),

```

### #2.3 Study interaction 
```{r}
library(survival)
fit1 <- coxph(Surv(time, event_1y) ~ Cluster + diseaseDuration , # + Age 
 data = df_survival)

fit2 <- coxph(Surv(time, event_1y) ~ Cluster*diseaseDuration , # + Age 
 data = df_survival)
 
anova(fit1,fit2,test="Chisq")
```
```{r}
library(survival)
fit1 <- coxph(Surv(time, event_1y) ~ Cluster + Sex , # + Age 
 data = df_survival)

fit2 <- coxph(Surv(time, event_1y) ~ Cluster*Sex , # + Age 
 data = df_survival)
 
anova(fit1,fit2,test="Chisq")
```

### #2.3 Calculate p-val for trend

```{r}
surv_pvalue(fitLC)
```

## #2.4 Check for remission
```{r}

# Import survival data (generate in #7.2.3 7_survival_analysis.ipynb)
df_survival <- as.data.frame(read_csv("new_data/5_clustering/df_survivaldata_mmae_Remission_setB.csv", col_types = cols(...1 = col_skip()))) # df_survivaldata_mmae_Remission_setB
head(df_survival)
```

### #2.2 Render Kaplan-Meier curve for 1 year

```{r}
library(survival)
library(survminer)
library("ggpubr")       # for ggarrange

df_survival$Cluster <- as.factor(df_survival$Cluster)
df_survival$time <- as.integer(df_survival$time_till_Rem) 
df_survival$event_Rem <- as.integer(df_survival$event_1y_Rem) 


fitLC  <- survfit(Surv(time, event_Rem) ~ Cluster, data = df_survival)

p1 <- ggsurvplot(fitLC
             , data = df_survival
             , risk.table = TRUE
             , conf.int=TRUE
             , pval = TRUE
             , break.x.by = 100
             , break.time.by= 100
             , xlab ="Time in Months"
             # , xlim = c(0, 60)        ## commented out !
             , ylab="% of patients still on initial MTX"
             , ggtheme = theme_minimal()
             , risk.table.y.text.col = T
             , risk.table.y.text = FALSE
             , risk.table.fontsize = 2
              , cumevents = TRUE
              , cumevents.fontsize =2   
             
             )

### save parts of the original graph    
surv_plot <- p1$plot
surv_table <- p1$table# +

### zoom in on the surv_plot    
surv_plot2 <- surv_plot + coord_cartesian(xlim = c(0,365), ylim = c(0.5,1.0))
surv_table <- surv_table + coord_cartesian(xlim = c(0,365))

### put it back together
ggarrange(surv_plot2, surv_table, ncol = 1, heights = c(3, 1))

# Save figure
ggsave("figures/5_treatment/survival_1y_Remission.png")
ggsave("figures/5_treatment/survival_1y_Remission.pdf")
#surv_plot2
#+ coord_cartesian(xlim=c(0, 365))#  xlim=c(0, 365),

```

### #2.6 Calculate p-val for trend w/ Remission

```{r}
surv_pvalue(fitLC)
```

## #3. Check for confounding physician

Import composite table with survival + metadata that we generated in notebook 9 (offshoots/9_Downstream_Treatment_analysis.ipynb) at step #2.3

```{r}
## Import EHR scaled data
library("readr")

df_ehr_scaled <- read_delim("new_data/offshoots/df_ehr_scaled_withOutcome_woCounts.csv", delim = "|", escape_double = FALSE, trim_ws = TRUE)
head(df_ehr_scaled)


df_ehr_remission  <- read_delim("new_data/offshoots/df_ehr_REMISSION.csv", delim = "|", escape_double = FALSE, trim_ws = TRUE)
head(df_ehr_remission)

```

```{r}
df_ehr_scaled$succes_1y <- +(!df_ehr_scaled$event_1y) 
```

### #3.1 Calculate the LISI score

We wanted to ensure that the clusters were not biased for certain physicians. Therefore, we rendered a LISI-density plot to show the diversity of physicians near the average patient , see figure x11. ('LISI' stands for Local Inverse Simpsons index)

```{r}
library(lisi)
library(ggplot2)
library(hrbrthemes)

Z <- subset(df_ehr_scaled, select=c(LF_0:LF_7)) 
meta_data <- subset(df_ehr_scaled, select=c(patnr:FinalTreatment, Physician))
meta_data$PHYSICIAN <- as.factor(meta_data$Physician) 
meta_data$Cluster <- as.factor(meta_data$Cluster) 

res <- compute_lisi(Z, meta_data, c('PHYSICIAN')) 
meta_data$PHYSICIAN_SCORE <- unlist(res)

# Visualize LISI distribution
p2 <- ggplot(data=meta_data, aes(x=PHYSICIAN_SCORE, group=Cluster, fill=Cluster)) +
    geom_density(adjust=1.5, alpha=.4) +
    theme_ipsum()

p2
```

## #4 Examine predictive ability of clusters for MTX response

### #4.1 Select MTX starters

```{r}

newdata <- as.data.frame(df_ehr_scaled[ which(df_ehr_scaled$`MTX-starter`==TRUE ),])
```

### #4.2 Create subsets for the different 'models'

```{r}

# MODEL A
X_A <- newdata[, c("RF_x", "aCCP_x", "Age_y", "Sex_x", "SJC", "TJC", "BSE_x",  "succes_1y")] 

# MODEL B
X_B <- newdata[, c("RF_x", "aCCP_x", "Age_y", "Sex_x", "SJC", "TJC", "BSE_x", "Cluster1", "Cluster2", "Cluster3", "Cluster4",  "succes_1y")] # "FEETcount", "HANDcount",

# MODEL C
X_C <- newdata[, c("RF_x", "aCCP_x", "Age_y", "Sex_x", "SJC", "TJC", "BSE_x", "FEETcount", "HANDcount",  "succes_1y")] # "FEETcount", "HANDcount

# MODEL D
X_D <- newdata[, c("RF_x", "aCCP_x", "Age_y", "Sex_x", "SJC", "TJC", "BSE_x", "FEETcount", "HANDcount", "Cluster1", "Cluster2", "Cluster3", "Cluster4",  "succes_1y")] # "FEETcount", "HANDcount


# Only cluster variables
X_clusterOnly <- newdata[, c("Cluster1", "Cluster2", "Cluster3", "Cluster4", "succes_1y")]

X_handOnly <- newdata[, c("HANDcount", "succes_1y")]

X_feetOnly <- newdata[, c("FEETcount", "succes_1y")]

# Only use aCCP as predictor
X_aCCPOnly <- newdata[, c("aCCP_x",  "succes_1y")]

# Only use ESR  as predictor
X_ESROnly <- newdata[, c("BSE_x", 'succes_1y')]

# Only use ESR  as predictor
X_SexOnly <- newdata[, c("Sex_x", 'succes_1y')]

# Only use DAS as predictor
X_DASOnly <- newdata[, c("DAS28(3)",  "succes_1y")]
```

### #4.3 Create the different predictive models

```{r}
#model
model_A <- glm(succes_1y ~ ., data = X_A, family = "binomial")
model_B <- glm(succes_1y ~ . , data = X_B, family = "binomial")
model_C <- glm(succes_1y ~ . , data = X_C, family = "binomial")
model_D <- glm(succes_1y ~ . , data = X_D, family = "binomial")

cluster_model <-  glm(succes_1y ~ ., data = X_clusterOnly, family = "binomial")
aCCP_model <-  glm(succes_1y ~ ., data = X_aCCPOnly, family = "binomial")
ESR_model <-  glm(succes_1y ~ ., data = X_ESROnly, family = "binomial")
Sex_model <-  glm(succes_1y ~ ., data = X_SexOnly, family = "binomial")
DAS_model <-  glm(succes_1y ~ ., data = X_DASOnly, family = "binomial")
null_model <- glm(succes_1y ~ 1, data = X, family = "binomial")

#hand_model <-  glm(succes_1y ~ ., data = X_handOnly, family = "binomial")
#feet_model <-  glm(succes_1y ~ ., data = X_feetOnly, family = "binomial")
```

### #4.4 Calculate additive value of the different models

```{r}
cat(paste("Calculate MacPhadden", "\n", 
      "Additive value of aCCP (Expl. var): ", 1-(logLik(aCCP_model)/logLik(null_model)), '\n',
      "Additive value of DAS (Expl. var): ", 1-(logLik(DAS_model)/logLik(null_model)), '\n',
      "Additive value of Sex (Expl. var): ", 1-(logLik(Sex_model)/logLik(null_model)), '\n',
      "Additive value of ESR (Expl. var): ", 1-(logLik(ESR_model)/logLik(null_model)), '\n',
      "Additive value of hand count (Expl. var): ", 1-(logLik(hand_model)/logLik(null_model)), '\n',
      "Additive value of feet count (Expl. var): ", 1-(logLik(feet_model)/logLik(null_model)), '\n',
      "Additive value of basemodel (Expl. var): ", 1-(logLik(base_model)/logLik(null_model)), '\n',
      "Additive value of original basemodel (Expl. var): ", 1-(logLik(base_model_old)/logLik(null_model)), '\n',
      "Additive value of clusterinfo alone (Expl. var): ", 1-(logLik(cluster_model)/logLik(null_model)), '\n',
      "Additive value of basemodel + cluster info (Expl. var): ", 1-(logLik(model)/logLik(null_model)), '\n',
      "Additive value of clusterinfo vs basemodel (Expl. var): ", 1-(logLik(model_B)/logLik(model_A))))
```

### #4.5 Calculate performance of model B

```{r}
library(broom)
require(MASS)
library(dplyr)

dt <- as.data.frame(exp(confint(model_B, level = 0.95))) # exp(coef(model))
dt$odds <- exp(coef(model_B))

# ALL
write.csv( tidy(model_B) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/coefs.csv")
write.csv( dt %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3), "results/treatmentResponse/odds.csv")
write.csv( glance(model_B ) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/an.csv")

summary(model_B)
```

### #4.5 Calculate performance of model C

```{r}
library(broom)
require(MASS)
library(dplyr)

dt <- as.data.frame(exp(confint(model_C, level = 0.95))) # exp(coef(model))
dt$odds <- exp(coef(model_C))

# ALL
write.csv( tidy(model_C) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/coefs.csv")
write.csv( dt %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3), "results/treatmentResponse/odds.csv")
write.csv( glance(model_C ) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/an.csv")

summary(model_C)
```

### #4.5 Calculate performance of model D
```{r}
library(broom)
require(MASS)
library(dplyr)

dt <- as.data.frame(exp(confint(model_D, level = 0.95))) # exp(coef(model))
dt$odds <- exp(coef(model_D))

# ALL
write.csv( tidy(model_D) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/coefs.csv")
write.csv( dt %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3), "results/treatmentResponse/odds.csv")
write.csv( glance(model_D ) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/an.csv")

summary(model_D)
```

### #4.6 Test if adding clusters to benchmark results in better model


```{r}
anova(base_model_old, model, test = "Chisq")

```
### #4.7 Check if hand & cluster model are similar

```{r}
anova(model_B, model_C, test = "Chisq")

```


```{r}
anova(hand_model, cluster_model, test = "Chisq")

```

### 4.8 Ensure association of clusters are not biased 
```{r}
library(broom)
require(MASS)
library(dplyr)
library(tidyverse)

X <- as.data.frame(df_survival[ which(df_survival$`MTX-starter`==TRUE ),])

X$succes_1y <- +(!X$event_1y)
clust_model <-  glm(succes_1y ~ Cluster1 + Cluster2 + Cluster3 + Cluster4, data = X, family = "binomial")
extra_model <- glm(succes_1y ~ MannTillTreat + SJC + TJC + aCCP + diseaseDuration + symptomDuration + Cluster1 + Cluster2 + Cluster3 + Cluster4, data = X, family = "binomial")

dt <- as.data.frame(exp(confint(extra_model, level = 0.95))) # exp(coef(model))
dt$odds <- exp(coef(extra_model))

# ALL
write.csv( tidy(extra_model) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/checkBias_coefs.csv")
write.csv( dt %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3), "results/treatmentResponse/checkBias_odds.csv")
write.csv( glance(extra_model ) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/checkBias_an.csv")


summary(extra_model)
```

### 5. Infer statistical significance at baseline between two populations

```{r}
library(readr)

df_lab <- read_delim("new_data/7_final/BaselinePopulation_TOTAL.csv", ";", escape_double = FALSE, trim_ws = TRUE)
df_lab
```


### 6. Fit remission model
#### 6.1 Select 
```{r}
newdata2 <- as.data.frame(df_ehr_remission[,])

# MODEL A
X_A <- newdata2[, c("RF", "aCCP", "Age", "Sex", "SJC", "TJC", "BSE",  "event_1y_Rem")] 

# MODEL B
X_B <- newdata2[, c("RF", "aCCP", "Age", "Sex", "SJC", "TJC", "BSE","Cluster1", "Cluster2", "Cluster3", "Cluster4",  "event_1y_Rem")] # "FEETcount", "HANDcount",

# Only clusters
X_clusterOnly <- newdata2[, c("Cluster1", "Cluster2", "Cluster3", "Cluster4", "event_1y_Rem")]

model_A <- glm(event_1y_Rem ~ ., data = X_A, family = "binomial")
model_B <- glm(event_1y_Rem ~ . , data = X_B, family = "binomial")
model_cluster <- glm(event_1y_Rem ~ . , data = X_clusterOnly, family = "binomial")

null_model <- glm(event_1y_Rem ~ 1, data = X_A, family = "binomial")

aCCP_model <- glm(event_1y_Rem ~ aCCP, data = X_A, family = "binomial")
RF_model <- glm(event_1y_Rem ~ RF, data = X_A, family = "binomial")
Age_model <- glm(event_1y_Rem ~ Age, data = X_A, family = "binomial")
ESR_model <- glm(event_1y_Rem ~ BSE, data = X_A, family = "binomial")
SJC_model <- glm(event_1y_Rem ~ SJC, data = X_A, family = "binomial")
TJC_model <- glm(event_1y_Rem ~ TJC, data = X_A, family = "binomial")
Sex_model <- glm(event_1y_Rem ~ Sex, data = X_A, family = "binomial")

cat(paste("Calculate MacPhadden", "\n", 
      "Additive value of model A (Expl. var): ", 1-(logLik(model_A)/logLik(null_model)), '\n',
      "Additive value of model B (Expl. var): ", 1-(logLik(model_B)/logLik(null_model)), '\n',
      "Additive value of RF (Expl. var): ", 1-(logLik(RF_model)/logLik(null_model)), '\n',
      "Additive value of aCCP (Expl. var): ", 1-(logLik(aCCP_model)/logLik(null_model)), '\n',
      "Additive value of Age (Expl. var): ", 1-(logLik(Age_model)/logLik(null_model)), '\n',
      "Additive value of BSE (Expl. var): ", 1-(logLik(ESR_model)/logLik(null_model)), '\n',
      "Additive value of SJC (Expl. var): ", 1-(logLik(SJC_model)/logLik(null_model)), '\n',
      "Additive value of TJC (Expl. var): ", 1-(logLik(TJC_model)/logLik(null_model)), '\n',
      "Additive value of Sex (Expl. var): ", 1-(logLik(Sex_model)/logLik(null_model)), '\n',
      "Additive value of clustermodel (Expl. var): ", 1-(logLik(model_cluster)/logLik(null_model))))

```

### #6.2 Calculate performance of model B

```{r}
library(broom)
require(MASS)
library(dplyr)

dt <- as.data.frame(exp(confint(model_B, level = 0.95))) # exp(coef(model))
dt$odds <- exp(coef(model_B))

# ALL
write.csv( tidy(model_B) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/coefs.csv")
write.csv( dt %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3), "results/treatmentResponse/odds.csv")
write.csv( glance(model_B ) %>% as.data.frame() %>%
  mutate(across(everything(), format, decimal.mark = ",")) %>% format(scientific=FALSE , digits = 3) , "results/treatmentResponse/an.csv")

summary(model_B)
```
### #6.3 Test if adding clusters to benchmark results in better model

```{r}
anova(model_B, model_A, test = "Chisq")
#anova(model_cluster, null_model, test = "Chisq")

```